---
title: "Coursera Data Products Class Project"
author: "Juan Chacin"
date: "October 23, 2017"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

- Company ACME maintains an inventory of pumps for sale
- These pumps are expensive and have a lead time from their supplier of 6 months
- The company would like to have enough pumps on hand to meet demand but not so many that they sit in inventory for months
- They have historical data for demand expressed as monthly sales average and standard deviation (normal distribution)
- They would like to use a Monte Carlo method to estimate when the probability of running out of pumps is greater than 10% in the next 6 months so they know when to order more

## Calculation method {.smaller}

- An R script was written to calculate scenarios based on historical data
- The inputs for the script are:
1. Average monthly pump usage
2. Standard deviation of monthly usage
3. Quantity on hand in inventory
- The script then performs 2000 (user input) simulations where:
1. A monthly usage is calculated for each of the next 6 months using a normal random distribution
2. That number is subtracted from the initial inventory
- At the end of the 2000 simulations, the probability of having zero pumps in inventory for each month is calculated 
- The result is presented in graphical form to the user for a purchasing decision

## Results {.smaller}
The results for a calculation are shown. The inputs were:

- Initial inventory of 55 pumps
- Monthly average usage of 10 pumps with a standard deviation of 3.5 pumps

```{r main code, echo = FALSE, fig.height=2.5}
# Clear the workspace and load libraries
rm(list = ls())

library("ggplot2",quietly = TRUE)

# Vector of dates expressed as calendar quarters

n_months <- 6       # Months into the future to estimate 
initial_date <- as.Date("10/1/2017","%m/%d/%Y")
time_ser <- as.Date(seq(initial_date,by="1 month",length.out=n_months), format = "%m%d%y")

# Monthly part use rate from past history. 
use_ave <- 10
use_std <- 3.5

# Start the simulation loop for statistical forecast
n_sim <- 20000                                    # Number of simulations

# Initialize needed arrays to hold simulation data
Invent <- matrix(nrow=(n_sim),ncol=n_months)       # matrix to hold inventory numbers for each simulation
Prob_Inven <- vector(length = n_months)             # Probability of running out. New systems

for (n in 1:n_sim){                 # Loop for number of simulations
    
    # Initial inventory values for the simulation
    Inven_ini <- 55                 # Initial quantity at hand in inventory

    # Loop for all quarters in the simulation 
    for (i in 1:n_months){              

        # Calculate usage for new systems using normal distribution with past averages
        use_month <- ceiling(rnorm(1,use_ave,use_std))
        
        if (use_month < 0) use_month <- 0     # Make sure the number is not less than zero
    
        Inven_ini <- Inven_ini - use_month
        Invent[n,i] <- Inven_ini              # Update inventory matrix
        
    }
}

# Compute quantiles for each months 

for (i in 1:n_months){
    P = ecdf(Invent[,i])
    Prob_Inven[i] <- P(0)*100
}

# Plot results

plot_data1 <- data.frame(time_ser,Prob_Inven)
names(plot_data1) <- c("Time","Inventory")

plot1 = ggplot(data = plot_data1) + geom_line(aes(x=plot_data1$Time,y=plot_data1$Inventory)) +  labs(title="Probability of Running Out by Date", x="Date",y="Probability [%]") + theme(plot.title = element_text(size = 18)) + theme(axis.title = element_text(size = 14))
```

```{r Plot results}
print(plot1)
```

## Summary

- A Monte Carlo calculation was performed to estimate the probability of running out of pumps in the next few months
- The result shows that the probability reaches 10% around early January, 2018
- Since that is inside of the 6 month lead time, new pumps should be ordered now from the supplier
- This document contains all the code needed to continue to update the results as new monthly data becomes available

